-------
Group Members: Gabriel Wright, Brennan Sauser
Project: Student Spending
Class: DS 202
output: github_document
-------

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

#Analysis on Student Spending Habits
## Introduction:
Goal of the project is to explore the dataset to be able to visualize student spending habits

Exploring the following questions/prompts:
1) What percentage is their monthly income is used for other expenses (non-school expenses)?

2) Does different majors affect student's monthly income?

3) Which category has the highest expenses?

4) Which Majors have the highest Tuitions

5) School Expenses based on Major

6) Does a student's gender have any indication of their major (what majors each gender typically go into)

7) Difference in student's expenses based on gender

8) student spending based on year

9) Major distribution by year and gender

10) Correlations of each variable/attribute

## Data Structure:
Dataset: "https://www.kaggle.com/datasets/sumanthnimmagadda/student-spending-dataset/data"
Variables Inlcuded in the dataset:
1) Age:
2) Gender:
3) Year in School:
4) Major:
5) Montly Income: of the student (in dollars)
6) Financial Aid: financial aid recieved by the student (in dollars)
7) Tuition: Expenses for tuition (in dollars)
8) Housing: Expenses for Housing (in dollars)
9) Food: Expenses for food (in dollars)
10) Transportation: Expenses for transportation
11) Book and Supplies: Expenses for books and supplies
12) Entertainment: Expenses for Entertainment (in dollars)
13) Personal Care: Expenses for personal care items (in dollars)
14) Technology: Expenses for technology (in dollars)
15) Health & Wellnesss: Expenses for health and wellness (in dollars)
16) Miscellaneous: Miscellaneous expenses (in dollars)
17) Preferred Payment Method: Method of payment (Cash, Credit/Debit card, Moblie Payment App)

## Cleaning
``` {r}
spending <- read_csv('student_spending.csv')
spending
```

``` {r}
#Checking for null values
sum(is.na(spending))
#No null or missing values

#Summary of data
summary(spending)

#Checking dimensions of the data
dim(spending)
#There are 1000 rows and 18 columns

#Removing Preferred Payment Method column since it is not relevant to the questions we want to answer
spending <- spending %>% select(-preferred_payment_method)

#Adding total expenses column and other expenses column (column that excludes tuition costs from its sum) for simplicity (for easy graphing)
spending <- spending %>% mutate(total_expenses = rowSums(pick(tuition:miscellaneous)), other_expenses = rowSums(pick(food, transportation, entertainment, personal_care, technology, health_wellness, miscellaneous)))
spending
```
## Statistical information to take note of when investigating the data
``` {r}
#Number of each gender
spending %>% group_by(gender) %>%
  summarize(n = n()) %>% ungroup()
#More male students in the data than the other two
```
``` {r}
#Number of each major
spending %>% group_by(major) %>%
  summarize(n = n()) %>% ungroup()
#Biology has the most amount of students in the program
```


## Results/Questions we want to answer
``` {r}
### Percentage of monthly income used for expenses other than school costs

```

``` {r}
### Distribution of majors based on their income
spending %>% group_by(major) %>% ggplot(aes(x = major, y = monthly_income, fill = major)) + geom_col()
#Biology students tend to make the most among college students
```

``` {r}
### Which category produces the highest expenses
spending %>% summarize() %>% ungroup()
```

``` {r}
### Which Majors have the highest tuition 

#Bar graph displaying the total expenses of each major
spending %>% group_by(major) %>% ggplot(aes(x = major, y = total_expenses)) + geom_col()
# Biology students tend to have the most expenses, with economic students with the second most expenses
```

``` {r}
### School Expenses based on major

```

``` {r}
### Distribution of Gender and Majors

#Graph of students gender and major
spending %>% group_by(major, gender) %>% ggplot(aes(x = major, fill = gender)) + geom_bar(position = position_dodge(preserve = 'single')) 
#Appears biology students have the most non-binary students. Economics and Biology are tied for having the most males in their programs
```

``` {r}
### Difference in students expenses and income based on gender

#Displays each genders average income
spending %>% group_by(gender) %>%
  summarize(avg_income = mean(monthly_income),
            n = n(), avg_expenses = mean(total_expenses)) %>% ungroup()
#Female and Non-binary people have an higher average of monthly income than men. However men on average tend to have the highest total expenses




#spending %>% ggplot(aes())
```

``` {r}
### Student spending based on year in school

#spending %>% ggplot(aes())
```

``` {r}
### Distribution of total different expenses 

```
