-------
Group Members: Gabriel Wright, Brennan Sauser
Project: Student Spending
Class: DS 202
output: github_document
-------

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(dplyr)
```

#Analysis on Student Spending Habits
## Introduction:
Goal of the project is to explore the dataset to be able to visualize student spending habits

Exploring the following questions/prompts:
1) What percentage is their monthly income is used for other expenses?
2) Does different majors affect student's income and expenses?

3) Which category has the highest expenses?

4) Which Majors have the highest Tuitions

5) School Expenses based on Major

6) Does a student's gender have any indication of their major (what majors each gender typically go into)

7) Difference in student's expenses based on gender

8) student spending based on year

9) Distribution of Methods for paying each expense

10) Major distribution (year, gender)

11) Correlations of each variable/attribute

## Data Structure:
Dataset: "https://www.kaggle.com/datasets/sumanthnimmagadda/student-spending-dataset/data"
Variables Inlcuded in the dataset:
1) Age:
2) Gender:
3) Year in School:
4) Major:
5) Montly Income: of the student (in dollars)
6) Financial Aid: financial aid recieved by the student (in dollars)
7) Tuition: Expenses for tuition (in dollars)
8) Housing: Expenses for Housing (in dollars)
9) Food: Expenses for food (in dollars)
10) Transportation: Expenses for transportation
11) Book and Supplies: Expenses for books and supplies
12) Entertainment: Expenses for Entertainment (in dollars)
13) Personal Care: Expenses for personal care items (in dollars)
14) Technology: Expenses for technology (in dollars)
15) Health & Wellnesss: Expenses for health and wellness (in dollars)
16) Miscellaneous: Miscellaneous expenses (in dollars)
17) Preferred Payment Method: Method of payment (Cash, Credit/Debit card, Moblie Payment App)

## Cleaning
``` {r}
```

